---
title: "Class11"
author: "Leyna Nguyen (PID:A15422197)"
date: "11/2/2021"
output: pdf_document
---
> Q1: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

87.55% (160736/183584) of the structures are solved by X-ray and 4.92% (9027/183584) of the structures are solved by electron microscopy. 

> Q2: What proportion of structures in the PDB are protein?

87.36% (160385/183584) of the structures are protein.

> Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There are 4411 HIV-1 protease structures.

> Q4: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?

Because the hydrogen atom and oxygen atoms that form a water molecule are so small, we only see the structure of a water molecule as 1 atom. 

> Q5: There is a conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have (see note below)?

This water molecule has residue number 308.

```{r}
library(bio3d)
pdb <- read.pdb("1hsg.pdb")
pdb
```

```{r}
attributes(pdb)
head(pdb$atom)
```






