---
title: "Class17"
author: "Leyna Nguyen (PID A15422197)"
date: "11/23/2021"
output: 
  pdf_document:
    latex_engine: xelatex
---

# 1. Getting Started

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
head(vax$as_of_date)
tail(vax$as_of_date)
```

> Q1. What column details the total number of people fully vaccinated?

The column called "persons_fully_vaccinated" details the total number of people fully vaccinated. 

> Q2. What column details the Zip code tabulation area?

The column called "zip_code_tabulation_area" details the zip code tabulation area. 

> Q3. What is the earliest date in this dataset?

The earliest date is 2021-01-05. 

> Q4. What is the latest date in this dataset?

The latest date is 2021-11-16.

```{r}
skimr::skim(vax)
```

> Q5. How many numeric columns are in this dataset?

```{r}
ncol(vax)
```
There are 9 numeric columns. 

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
sum( is.na(vax$persons_fully_vaccinated) )
```
There are 8256 NA values in the persons_fully_vaccinated column.

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

```{r}
sum(vax$persons_fully_vaccinated)
```

We will use the **lubridate** package to make life a lot easier when dealing with dates and times: 

```{r}
library(lubridate)
```

```{r}
today()
```
Now I can do useful math with dates more easily:
```{r}
# Speciffy that we are using the Year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
today() - vax$as_of_date[1]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

> Q9. How many days have passed since the last update of the dataset?

```{r}
today() - vax$as_of_date[nrow(vax)]
```
7 days have passed since the last update of the dataset.

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))


46*7
```
There are 46 unique dates in the dataset.

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

# Pull data for all ZIP codes in the dataset
#zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )

```{r}
# Subset to San Diego county only areas
sd <- vax[vax$county == "San Diego", ]
```

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

> Q11. How many distinct zip codes are listed for San Diego County?

```{r}
length(unique((sd$zip_code_tabulation_area)))

```
There are 107 distinct zip codes listed for San Diego county. 

> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
ind <- which.max(sd$age12_plus_population)
sd[ind,]
```
The zip code area 92154 has the largest 12+ population in this dataset.

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?

```{r}
sd.now <- filter(sd, as_of_date=="2021-11-09")
mean(sd.now$percent_of_population_fully_vaccinated, na.rm=TRUE)
```
The overall average is 67.3%. 

> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```
```{r}
library(ggplot2)

ggplot(sd.now) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram(bins=15)

```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
filter(sd.now, zip_code_tabulation_area=="92037")

ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

```{r}
# Subset to all CA areas with a population as large as 92037
ucsd[1,]$age5_plus_population
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

#head(vax.36)
```

> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  geom_hline(yintercept=0.67276, color="red") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
summary(vax.36)
```
gasjfladfvkjerhgdwtkejhrgk

> Q18. Using ggplot generate a histogram of this data.

```{r}
library(ggplot2)

ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram() +
  labs(x="Percent Vaccinated", y="Count")
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
The zip code 92040 is below the average and the zip code 92109 is above the average.

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)

ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with a population above 36k are shown") +
  geom_hline(yintercept=0.66, col="red")
```

> Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

I'm staying in San Diego for the break, but since I can imagine that a lot of people will be going home for the break, I think I'd feel safer if class wasn't in-person. 










