---
title: "Class10"
author: "Leyna Nguyen (PID: A15422197)"
date: "10/28/2021"
output: html_document
---

# 1. Importing candy data

```{r}
candy_file <- "candy-data.csv"
candy = read.csv(candy_file, row.names=1)
head(candy)
```
> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```
There are 85 different candy types.

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```
There are 38 fruity candy types.

# 2. What is your favorite candy?

```{r}
# How many people prefer Twix over a randomly chosen candy?
candy["Twix", ]$winpercent
```

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
# My favorite candy is Kit Kat! 
candy["Kit Kat", ]$winpercent 
```
Kit Kat's winpercent value is 76.7686%. 

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent 
```
Kit Kat's winpercent value is 76.7686%. 

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
A Tootsie Roll Snack Bars' winpercent value is 49.6535%. 

```{r}
library("skimr")
skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, winpercent is on a different scale than the majority of other columns. 

> Q7. What do you think a zero and one represent for the candy$chocolate column?

```{r}

```

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent, main="Histogram of Winpercent", xlab="Winpercent", )
```


> Q9. Is the distribution of winpercent values symmetrical?

No, it's not. The distribution is skewed to the right???

> Q10. Is the center of the distribution above or below 50%?

The center of the distribution is below 50%. 

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

First we need to find all the chocolate candy rows in the `candy` data. 
```{r}
mean(candy$winpercent[as.logical(candy$chocolate)])
mean(candy$winpercent[as.logical(candy$fruity)])

inds <- as.logical(candy$chocolate)
candy[inds,]$winpercent
```
On average, chocolate candy is ranked higher than fruity candy. 

> Q12. Is this difference statistically significant?

```{r}
t.test(candy$winpercent[as.logical(candy$chocolate)], candy$winpercent[as.logical(candy$fruity)])
```
Yes, this difference is statistically significant because the p-value is lower than 0.05. 

# 3. Overall candy rankings

> Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```
The 5 least liked candy types are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters. 

```{r}
candy %>% 
  arrange (winpercent) %>% 
  head(5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n=5)
decreasing=TRUE before n=5
```
The top 5 favorite candy types are Snickers, Kit Kat, Twix, Reese's Miniatures, and Reese's Peanut Butter Cups. 

> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent.

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) + 
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols)
```

> Q17. What is the worst ranked chocolate candy?

The worst ranked chocolate candy is Sixlets.

> Q18. What is the best ranked fruity candy?

The best ranked fruity candy is Starburst. 

# 4. Taking a look at pricepercent

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reese's Miniatures

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
tail( candy[ord,c(11,12)], n=5 )
```

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
The top 5 most expensive candy types are Nik L Nip, Nestle Smarties, Ring Pop, Hershey's Krackel, and Hershey's Milk Chocolate. Of these, Nik L Nip is the least popular. 

> Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chart” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

```{r}
# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

# 5. Exploring the correlation structure

```{r}
install.packages('corrplot')
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

The variables chocolate and fruity are anti-correlated.

> Q23. Similarly, what two variables are most positively correlated?

The variables chocolate and bar are most positively correlated.

# 6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[, 1:2])

# Add some color
plot(pca$x[,1:2], col=my_cols, pch=16)
```
Let's use ggplot to make a nicer version of this. Here we need a new data.frame that contains all the info I want displayed in my plot so I want to combine my PCA results with all the candy data. 
```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])

# Plot a new ggplot
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
        geom_text_repel(col=my_cols, size=3, max.overlaps=)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
```{r}
install.packages('plotly')
library(plotly)
ggplotly(p)
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

The variables fruity, pluribus, and hard are picked up strongly by PC1 in the positive direction. This makes sense because  










